### 2024-05-31 thesis/appendix.tex
May 31st 2024

\begin{table}
\caption{Pinball loss for Boston housing data}
\begin{tabular}{lllll}
\toprule
    & Linear qr & Gbm qr & Quantile forest & KQR \\
\midrule
CRPS & 1.3785678 & 1.1418540 & 1.0587686 & \textbf{1.0297572} \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}
    \caption{Pinball loss quantile-wise for Boston data}
\begin{tabular}{lllll}
\toprule
    & Linear qr & Gbm qr & Quantile forest & KQR \\
\midrule
0.1 & 0.729749 & 0.771714 & 0.588441 & \textbf{0.578898} \\
0.2 & 1.122582 & 1.033442 & 0.932824 & \textbf{0.869145} \\
0.3 & 1.479486 & 1.170642 & 1.153765 & \textbf{1.142783} \\
0.4 & 1.712577 & 1.436263 & 1.352667 & \textbf{1.331955} \\
0.5 & 1.911385 & \textbf{1.344361} & 1.408333 & 1.396300 \\
0.6 & 1.989514 & 1.448885 & 1.464902 & \textbf{1.431705} \\
0.7 & 1.938362 & 1.508741 & 1.427912 & \textbf{1.382772} \\
0.8 & 1.658058 & 1.497901 & 1.275059 & \textbf{1.245288} \\
0.9 & 1.243965 & 1.206591 & 0.983784 & \textbf{0.918725} \\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{table}
    \caption{Pinball loss for Abalone data}
\begin{tabular}{lllll}
    \toprule
     & Linear qr & Gbm qr & Quantile forest & KQR \\
    \midrule
    CRPS & 0.5613975 & 0.5531938 & \textbf{0.5212990} & 0.5252491 \\
    \bottomrule
    \end{tabular}
\end{table}

\begin{table}
    \caption{Pinball loss quantile-wise for Abalone data}
    \begin{tabular}{lllll}
    \toprule
     & Linear qr & Gbm qr & Quantile forest & KQR \\
    \midrule
    0.1 & 0.277903 & 0.290531 & 0.274079 & \textbf{0.269287} \\
    0.2 & 0.469361 & 0.488079 & \textbf{0.453110} & 0.457286 \\
    0.3 & 0.621961 & 0.625633 & \textbf{0.580766} & 0.596791 \\
    0.4 & 0.729757 & 0.715875 & \textbf{0.689904} & 0.691310 \\
    0.5 & 0.794695 & 0.766185 & \textbf{0.735945} & 0.740834 \\
    0.6 & 0.810691 & 0.785769 & \textbf{0.744928} & 0.746636 \\
    0.7 & 0.769587 & 0.730318 & \textbf{0.700287} & 0.710392 \\
    0.8 & 0.667776 & 0.656913 & 0.609378 & \textbf{0.608334} \\
    0.9 & 0.472244 & 0.472635 & \textbf{0.424593} & 0.431621 \\
    \bottomrule
    \end{tabular}
\end{table}

@@

\begin{table}
\caption{Pinball loss for Vehicle data}
\begin{tabular}{lllll}
    \toprule
     & Linear qr & Gbm qr & Quantile forest & KQR \\
    \midrule
    CRPS & 0.4054449 & 0.2289554 & 0.2844410 & \textbf{0.2204343} \\
    \bottomrule
    \end{tabular}
\end{table}

\begin{table}
    \caption{Pinball loss quantile-wise for Vehicle data}
    \begin{tabular}{lllll}
    \toprule
     & Linear qr & Gbm qr & Quantile forest & KQR \\
    \midrule
    0.1 & 0.254649 & \textbf{0.139849} & 0.170489 & 0.182490 \\
    0.2 & 0.403772 & 0.236339 & 0.285875 & \textbf{0.223165} \\
    0.3 & 0.548820 & 0.244086 & 0.357375 & \textbf{0.242963} \\
    0.4 & 0.576918 & 0.263169 & 0.389305 & \textbf{0.262835} \\
    0.5 & 0.554367 & 0.306123 & 0.410738 & \textbf{0.295878} \\
    0.6 & 0.563046 & 0.335363 & 0.398125 & \textbf{0.306062} \\
    0.7 & 0.516019 & 0.287490 & 0.326572 & \textbf{0.283716} \\
    0.8 & 0.407742 & 0.261882 & 0.313698 & \textbf{0.235793} \\
    0.9 & 0.229115 & 0.215253 & 0.192233 & \textbf{0.171440} \\
    \bottomrule
    \end{tabular}
\end{table}

>>>>

\begin{table}
\caption{Continous ranked probability score for Boston housing data}
\begin{tabular}{lllll}
\toprule
    & Linear qr & Gbm qr & Quantile forest & KQR \\
\midrule
CRPS & 1.3785678 & 1.1418540 & 1.0587686 & \textbf{1.0297572} \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}
    \caption{Pinball loss quantile-wise for Boston data}
\begin{tabular}{lllll}
\toprule
  Quantile  & Linear qr & Gbm qr & Quantile forest & KQR \\
\midrule
0.1 & 0.729749 & 0.771714 & 0.588441 & \textbf{0.578898} \\
0.2 & 1.122582 & 1.033442 & 0.932824 & \textbf{0.869145} \\
0.3 & 1.479486 & 1.170642 & 1.153765 & \textbf{1.142783} \\
0.4 & 1.712577 & 1.436263 & 1.352667 & \textbf{1.331955} \\
0.5 & 1.911385 & \textbf{1.344361} & 1.408333 & 1.396300 \\
0.6 & 1.989514 & 1.448885 & 1.464902 & \textbf{1.431705} \\
0.7 & 1.938362 & 1.508741 & 1.427912 & \textbf{1.382772} \\
0.8 & 1.658058 & 1.497901 & 1.275059 & \textbf{1.245288} \\
0.9 & 1.243965 & 1.206591 & 0.983784 & \textbf{0.918725} \\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{table}
    \caption{Continous ranked probability score for Abalone data}
\begin{tabular}{lllll}
    \toprule
  & Linear qr & Gbm qr & Quantile forest & KQR \\
    \midrule
    CRPS & 0.5613975 & 0.5531938 & \textbf{0.5212990} & 0.5252491 \\
    \bottomrule
    \end{tabular}
\end{table}

\begin{table}
    \caption{Pinball loss quantile-wise for Abalone data}
    \begin{tabular}{lllll}
    \toprule
    Quantile & Linear qr & Gbm qr & Quantile forest & KQR \\
    \midrule
    0.1 & 0.277903 & 0.290531 & 0.274079 & \textbf{0.269287} \\
    0.2 & 0.469361 & 0.488079 & \textbf{0.453110} & 0.457286 \\
    0.3 & 0.621961 & 0.625633 & \textbf{0.580766} & 0.596791 \\
    0.4 & 0.729757 & 0.715875 & \textbf{0.689904} & 0.691310 \\
    0.5 & 0.794695 & 0.766185 & \textbf{0.735945} & 0.740834 \\
    0.6 & 0.810691 & 0.785769 & \textbf{0.744928} & 0.746636 \\
    0.7 & 0.769587 & 0.730318 & \textbf{0.700287} & 0.710392 \\
    0.8 & 0.667776 & 0.656913 & 0.609378 & \textbf{0.608334} \\
    0.9 & 0.472244 & 0.472635 & \textbf{0.424593} & 0.431621 \\
    \bottomrule
    \end{tabular}
\end{table}

@@

\begin{table}
\caption{Continous ranked probability score for Vehicle data}
\begin{tabular}{lllll}
    \toprule
     & Linear qr & Gbm qr & Quantile forest & KQR \\
    \midrule
    CRPS & 0.4054449 & 0.2289554 & 0.2844410 & \textbf{0.2204343} \\
    \bottomrule
    \end{tabular}
\end{table}

\begin{table}
    \caption{Pinball loss quantile-wise for Vehicle data}
    \begin{tabular}{lllll}
    \toprule
    Quantile & Linear qr & Gbm qr & Quantile forest & KQR \\
    \midrule
    0.1 & 0.254649 & \textbf{0.139849} & 0.170489 & 0.182490 \\
    0.2 & 0.403772 & 0.236339 & 0.285875 & \textbf{0.223165} \\
    0.3 & 0.548820 & 0.244086 & 0.357375 & \textbf{0.242963} \\
    0.4 & 0.576918 & 0.263169 & 0.389305 & \textbf{0.262835} \\
    0.5 & 0.554367 & 0.306123 & 0.410738 & \textbf{0.295878} \\
    0.6 & 0.563046 & 0.335363 & 0.398125 & \textbf{0.306062} \\
    0.7 & 0.516019 & 0.287490 & 0.326572 & \textbf{0.283716} \\
    0.8 & 0.407742 & 0.261882 & 0.313698 & \textbf{0.235793} \\
    0.9 & 0.229115 & 0.215253 & 0.192233 & \textbf{0.171440} \\
    \bottomrule
    \end{tabular}
\end{table}

### 2024-05-31 thesis/input_files/acknowledgements.tex
May 31st 2024

% Thank:
% - supervisor, co
First and foremost I would like to express my deepest gratitude to my advisor and co-advisor, Prof. Multerer and Dr. Baroli for their invaluable feedback, continuous support and patience during my Master's Thesis. Furthermore, I thank them for having given me the opportunity to collaborate through this thesis work.
\\
% - master director
I could not have undertaken this journey without my Master directors Prof. Schenk and Prof. Wit. I am extremely grateful for their guidance, mentorship and advice throughout my Master studies.
\\
% - professors of faculty
% From Polya How to solve it
% forster independent thinking  by challenging curiosity, stimulating questions and delving into the motives behind theory.
% tutti i professori che ho avuto, shaped my interest, saziato sete curiosità
I would like to extend my deepest appreciation to all professors I had during my studies at Usi. Their stimulating questions and push to enquire motives behind theory challenged my curiosity, shaped my interests and fostered my independent thinking.
\\
% - research group 
% discussions, suggestions
% listening the challenges I faced and insightful suggestions
I also had great pleasure of discussing with the research group of Prof. Multerer the challenges I faced during this thesis work. I thank them for their affability  and insightful suggestions.
\\
% - classmates
% friends I made along the way
% mutual support and good times we had together
% be it during a group revision session before an exam of during light hearted moments.
Special thanks to my cohort members and to the friends I made along the way. I will never forget the good times we had together.
\clearpage

>>>>

% Thank:
% - supervisor, co
First and foremost I would like to express my deepest gratitude to my advisor and co-advisor, Prof.\ Multerer and Dr.\ Baroli for their invaluable feedback, continuous support and patience during my Master's Thesis. Furthermore, I thank them for having given me the opportunity to collaborate through this thesis work.
\\
% - master director
I could not have undertaken this journey without my Master directors Prof.\ Schenk and Prof.\ Wit. I am extremely grateful for their guidance, mentorship and advice throughout my Master studies.
\\
% - professors of faculty
% From Polya How to solve it
% forster independent thinking  by challenging curiosity, stimulating questions and delving into the motives behind theory.
% tutti i professori che ho avuto, shaped my interest, saziato sete curiosità
I would like to extend my deepest appreciation to all professors I had during my studies at Usi. Their stimulating questions and push to enquire motives behind theory challenged my curiosity, shaped my interests and fostered my independent thinking.
\\
% - research group 
% discussions, suggestions
% listening the challenges I faced and insightful suggestions
I also had great pleasure of discussing with the research group of Prof.\ Multerer the challenges I faced during this thesis work. I thank them for their affability  and insightful suggestions.
\\
% - classmates
% friends I made along the way
% mutual support and good times we had together
% be it during a group revision session before an exam of during light hearted moments.
Special thanks to my cohort members and to the friends I made along the way. I will never forget the good times we had together.
\clearpage

### 2024-05-31 thesis/input_files/analysis.tex
May 31st 2024

\begin{table}[!ht]
    \centering
\caption{Pinball loss kernel comparison for load in Austria (2021)}
\label{tab:secures_met_at}
\begin{adjustbox}{width=\textwidth}
    \begin{tabular}{lllllllllll}
        \toprule
        Quantile & \begin{tabular}[c]{@{}l@{}}Absolute\\ Laplacian\end{tabular} & \begin{tabular}[c]{@{}l@{}}Matern 0.5/\\ Laplacian\end{tabular} & Matern 1.5 & Matern 2.5 & \begin{tabular}[c]{@{}l@{}}Matern $\infty$/\\ Gaussian RBF\end{tabular} & Linear & Periodic & Polynomial & Sigmoid & Cosine \\
        \midrule
        0.1 & \textbf{0.024365} & 0.024592 & 0.024550 & 0.024611 & 0.024831 & 0.026853 & 0.025744 & 0.026004 & 0.027672 & 0.027816 \\
        0.2 & \textbf{0.040971} & 0.041406 & 0.041413 & 0.041487 & 0.041857 & 0.045007 & 0.043247 & 0.043720 & 0.045899 & 0.045879 \\
        0.3 & \textbf{0.052316} & 0.052969 & 0.052888 & 0.052973 & 0.053449 & 0.057880 & 0.055231 & 0.055446 & 0.058572 & 0.058454 \\
        0.4 & \textbf{0.058286} & 0.059018 & 0.058916 & 0.059078 & 0.059592 & 0.065770 & 0.061844 & 0.062495 & 0.066599 & 0.066335 \\
        0.5 & 0.060344 & 0.061100 & 0.061319 & 0.061541 & 0.062338 & 0.068197 & 0.063864 & 0.071731 & 0.069761 & 0.069574 \\
        0.6 & \textbf{0.058724} & 0.059507 & 0.059674 & 0.059930 & 0.060709 & 0.065828 & 0.062265 & 0.070269 & 0.067179 & 0.066970 \\
        0.7 & \textbf{0.053148} & 0.054189 & 0.054249 & 0.054372 & 0.054916 & 0.058548 & 0.055830 & 0.063015 & 0.059807 & 0.059419 \\
        0.8 & \textbf{0.042740} & 0.043368 & 0.043408 & 0.043510 & 0.043881 & 0.046017 & 0.044384 & 0.049813 & 0.047078 & 0.046662 \\
        0.9 & \textbf{0.026430} & 0.026583 & 0.026781 & 0.026817 & 0.026908 & 0.027050 & 0.026471 & 0.030279 & 0.028031 & 0.027986 \\
        \midrule
        CRPS & \textbf{0.046369} & 0.046970 & 0.047022 & 0.047146 & 0.047609 & 0.051239 & 0.048764 & 0.052530 & 0.052289 & 0.052122 \\
        \bottomrule
    \end{tabular}
\end{adjustbox}
\end{table}
From our tables, it can be concluded that the absolute Laplacian kernel is the one fitting best to the characteristics of load data.
Additionally, we can conclude that our results are in line with the ones published in \cite{he2017short}. In that study, similarly to what we did in this subsection, the authors carried out a comparitive study concerning the choice of kernels in KQR. Nevertheless notice, that the set of kernel we considered in our study is larger than the one presented in \cite{he2017short}.
Their paper showed evidence of the superiority of the Gaussian RBF among their set of kernel functions. Analogously, we can see from Tables \ref{tab:secures_met_de}, \ref{tab:secures_met_ch} and \ref{tab:secures_met_at} that the Matern kernels are the best in terms of pinball loss.
Finally, given that the features dimension is greater than one we have that the Absolute Laplacian and the Matern 0.5 kernels differ. The former kernel is better than the latter and this comes from their distance function. This is because the Manhattan distance is more robust than the Euclidean distance in terms of measuring similarity between objects as data grows in dimensionality \cite{Aggarwal2001}. %(curse of dimensionality)
Those kernel matrices for the Swiss data are compared in Figure \ref{fig:k_matrix_laplacian_comp}.

@@

Our results for this track are reported in Table \ref{tab:pinball loss gefcom2014 price data}. In this track, the top entries came from the teams: Tololo, Team Poland, GMD, and C3 Green Team; for a breakdown of each method's attributes, see \cite[Table 8]{hong2016probabilistic}. Finally, the Gaussian RBF probabilistic prediction for the $4^{th}$ July $2013$ zonal price at the $90\%$ confidence interval is visualized in Figure \ref{fig:price_task_4_gaussian_rbf}.

>>>>

\begin{table}[!ht]
    \centering
\caption{Pinball loss kernel comparison for load in Austria (2021)}
\label{tab:secures_met_at}
\begin{adjustbox}{width=\textwidth}
    \begin{tabular}{lllllllllll}
        \toprule
        Quantile & \begin{tabular}[c]{@{}l@{}}Absolute\\ Laplacian\end{tabular} & \begin{tabular}[c]{@{}l@{}}Matern 0.5/\\ Laplacian\end{tabular} & Matern 1.5 & Matern 2.5 & \begin{tabular}[c]{@{}l@{}}Matern $\infty$/\\ Gaussian RBF\end{tabular} & Linear & Periodic & Polynomial & Sigmoid & Cosine \\
        \midrule
        0.1 & \textbf{0.024365} & 0.024592 & 0.024550 & 0.024611 & 0.024831 & 0.026853 & 0.025744 & 0.026004 & 0.027672 & 0.027816 \\
        0.2 & \textbf{0.040971} & 0.041406 & 0.041413 & 0.041487 & 0.041857 & 0.045007 & 0.043247 & 0.043720 & 0.045899 & 0.045879 \\
        0.3 & \textbf{0.052316} & 0.052969 & 0.052888 & 0.052973 & 0.053449 & 0.057880 & 0.055231 & 0.055446 & 0.058572 & 0.058454 \\
        0.4 & \textbf{0.058286} & 0.059018 & 0.058916 & 0.059078 & 0.059592 & 0.065770 & 0.061844 & 0.062495 & 0.066599 & 0.066335 \\
        0.5 & 0.060344 & 0.061100 & 0.061319 & 0.061541 & 0.062338 & 0.068197 & 0.063864 & 0.071731 & 0.069761 & 0.069574 \\
        0.6 & \textbf{0.058724} & 0.059507 & 0.059674 & 0.059930 & 0.060709 & 0.065828 & 0.062265 & 0.070269 & 0.067179 & 0.066970 \\
        0.7 & \textbf{0.053148} & 0.054189 & 0.054249 & 0.054372 & 0.054916 & 0.058548 & 0.055830 & 0.063015 & 0.059807 & 0.059419 \\
        0.8 & \textbf{0.042740} & 0.043368 & 0.043408 & 0.043510 & 0.043881 & 0.046017 & 0.044384 & 0.049813 & 0.047078 & 0.046662 \\
        0.9 & \textbf{0.026430} & 0.026583 & 0.026781 & 0.026817 & 0.026908 & 0.027050 & 0.026471 & 0.030279 & 0.028031 & 0.027986 \\
        \midrule
        CRPS & \textbf{0.046369} & 0.046970 & 0.047022 & 0.047146 & 0.047609 & 0.051239 & 0.048764 & 0.052530 & 0.052289 & 0.052122 \\
        \bottomrule
    \end{tabular}
\end{adjustbox}
\end{table}
From our tables, it can be concluded that the absolute Laplacian kernel is the one fitting best to the characteristics of load data.
Additionally, we can conclude that our results are in line with the ones published in \cite{he2017short}. In that study, similarly to what we did in this subsection, the authors carried out a comparitive study concerning the choice of kernels in KQR. Nevertheless notice, that the set of kernel we considered in our study is larger than the one presented in \cite{he2017short}.
Their paper showed evidence of the superiority of the Gaussian RBF among their set of kernel functions. Analogously, we can see from Tables \ref{tab:secures_met_de}, \ref{tab:secures_met_ch} and \ref{tab:secures_met_at} that the Matern kernels are the best in terms of pinball loss.
The Periodic equipped with period $p=24$ shows good performance. 
After those kernels, we have the Linear and Polynomial kernel.
The worst among all kernels considered are the Cosine and the Sigmoid.
% , which is better suited for binary classification tasks. 
Finally, given that the features dimension is greater than one we have that the Absolute Laplacian and the Matern 0.5 kernels differ. The former kernel is better than the latter and this comes from their distance function. This is because the Manhattan distance is more robust than the Euclidean distance in terms of measuring similarity between objects as data grows in dimensionality \cite{Aggarwal2001}. %(curse of dimensionality)
Those kernel matrices for the Swiss data are compared in Figure \ref{fig:k_matrix_laplacian_comp}.

@@

Our results for this track are reported in Table \ref{tab:pinball loss gefcom2014 price data}. In this track, the top entries came from the teams: Tololo, Team Poland, GMD, and C3 Green Team; for a breakdown of each method's attributes, see \cite[Table 8]{hong2016probabilistic}. 
The Gaussian RBF probabilistic prediction for the $4^{th}$ July $2013$ zonal price at the $90\%$ confidence interval is visualized in Figure \ref{fig:price_task_4_gaussian_rbf}.
Cross validation results are visualised in Figure.

\begin{figure}
    \includegraphics[width=\textwidth]{images/cv_scatter.png}
    \label{fig:cv_scatter}
    \caption{Cross validation Gaussian RBF kernel task 4 GEFCom2014 price}
\end{figure}

### 2024-05-31 thesis/input_files/implementation.tex
May 31st 2024

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of \href{https://www.statsmodels.org/dev/generated/statsmodels.regression.quantile_regression.QuantReg.html}{quantile\_regression.QuantReg} from the regression module of statsmodels has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of \href{https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.GradientBoostingRegressor.html}{GradientBoostingRegressor} from the sklearn.ensemble submodule has been used.
\subsection{Quantile forest}
The implementation of \href{https://pypi.org/project/quantile-forest/}{quantile\_forest.RandomForestQuantileRegressor} has been used. This estimator is compatible with scikit-learn API \cite{Johnson2024}.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented Python open source library, thus the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialized by providing a quantile, the regularisation term $C$, the kernel family and its corresponding hyperparameters.
\\
In the fit method, we set up and solve the convex optimisation problem through the interior point method. This algorithm is taken from the cvxopt library, see its official manual \citeW{vandenberghe2010cvxopt} for a reference.
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the $Q$ matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them and provide them as a single matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix $X\_eval$ of independent variables, next we compute the kernel matrix between $X\_train$ and $X\_eval$ and obtain $y\_eval$ with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with in built scikit-learn methods like gridsearch, crossvalidation and scoring rules. Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}. Functions supported by our kernel quantile regression include: Gaussian RBF, absolute Laplacian, Matern 3/2, Matern 5/2, Linear, Cosine, Sigmoid, Periodic, Polynomial and custom composition of kernels.
% NOTICE, these kernel implementations differ in the form of passed parameters in the former we pass in the numerator gamma while in the latter we pass l, which stands for length scale in the denominator. To cross validate overt the same range of parameter across the different kernels we made sura that gamma=1/l.

>>>>

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of \href{https://www.statsmodels.org/dev/generated/statsmodels.regression.quantile_regression.QuantReg.html}{quantile\_regression.QuantReg} from the regression module of statsmodels has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of \href{https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.GradientBoostingRegressor.html}{GradientBoostingRegressor} from the sklearn.ensemble submodule has been used.
\subsection{Quantile forest}
The implementation of \href{https://pypi.org/project/quantile-forest/}{quantile\_forest.RandomForestQuantileRegressor} has been used. This estimator is compatible with scikit-learn API \cite{Johnson2024}.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented Python open source library, thus the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialized by providing a quantile, the regularisation term $C$, the kernel family and its corresponding hyperparameters.
\\
In the fit method, we set up and solve the convex optimisation problem through the interior point method. This algorithm is taken from the cvxopt library, see its official manual \citeW{vandenberghe2010cvxopt} for a reference.
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the $Q$ matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them and provide them as a single matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix $X\_eval$ of independent variables, next we compute the kernel matrix between $X\_train$ and $X\_eval$ and obtain $y\_eval$ with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with in built scikit-learn methods like gridsearch, crossvalidation and scoring rules. Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}. Functions supported by our kernel quantile regression include: Gaussian RBF, absolute Laplacian, Matern 3/2, Matern 5/2, Linear, Cosine, Sigmoid, Periodic, Polynomial and custom composition of kernels. 
Kernels from these two submodules use a different convention in the hyperparameters. The former has the lenghtscale parameter in the numerator while the latter in the denominator. In order to meaningfully compare across different kernels we made them consistent in this aspect. For this purpose we adhere to the convention of having lenghtscale hyperparameters in the denominator.

% NOTICE, these kernel implementations differ in the form of passed parameters in the former we pass in the numerator gamma while in the latter we pass l, which stands for length scale in the denominator. To cross validate overt the same range of parameter across the different kernels we made sura that gamma=1/l.

### 2024-05-31 thesis/input_files/paper1.tex
May 31st 2024

\section{Kernel families}
The selection of the kernel determines the class of functions that will be searched by the learning algorithm. Prior knowledge of the problem domain can help restricting the candidate families of kernel functions. 
Table \ref{tab:kernel types} contains popular kernel families in literature and applications.
\begin{table}
    \caption{Kernel types}
    \resizebox{\textwidth}{!}{
    \begin{tabular}{lll}
        \toprule
       Kernel function & Equation & Hyperparameters \\
       \midrule
       Linear &  $k(x,y)=xy$ &   \\
       Polynomial &  $k(x,y)=(x^\intercal y+c)^d$ &   c, d\\
       Gaussian RBF &  $k(x,y)=\exp\left(-\frac{\|x-y\|_{2}^2}{2\l^2}\right)$ &   $l$\\
       Exponential RBF/Laplacian& $k(x,y)=\exp\left(-\frac{\|x-y\|_{2}}{l}\right)$ &   $l$\\
       Absolute Laplacian& $k(x,y)=\exp\left(-\frac{\|x-y\|_{1}}{l}\right)$ &   $l$\\
       Hyperbolic/Sigmoid Kernel &  $k(x,y2)=\tanh(\gamma x^\intercal y+r)$ &  $\gamma$, r \\
       Periodic &  $k(x,y)=\exp\left(\frac{-2 sin^2\left(\frac{\pi}{p}\|x-y\|_{2} \right)}{l^2}\right) $ &   $p, l$\\
       Chi-squared kernel &  $k(x,y)=\exp\left(-\gamma \sum\limits_i \frac{(x_i - y_i)^2}{x_i+y_i}\right) $ &   $\gamma$\\
        Cosine  &  $k(x,y)=\frac{xy^\intercal}{\|x\|_{2}\|y\|_{2}}$ &   \\
       Matérn  &  $k(x, y) =  \frac{1}{\Gamma(\nu)2^{\nu-1}}\Bigg(
        \frac{\sqrt{2\nu}}{l} \|x-y|\|_{2}
        \Bigg)^\nu Y_\nu\Bigg(
        \frac{\sqrt{2\nu}}{l} \|x-y|\|_{2}\Bigg)$ &  $l, \nu$ \\

@@

When $\nu=\frac{1}{2}$ the Matérn kernel corresponds to the Laplacian kernel.
Furthermore, as $\nu \to \infty$ we have that the Matérn kernel converges to the Gaussian RBF kernel.
$k(x,y)=\exp\left(\frac{-\|x-y\|_{2}^2}{2l^2}\right)$

>>>>

\section{Kernel families}\label{kernel families}
The selection of the kernel determines the class of functions that will be searched by the learning algorithm. Prior knowledge of the problem domain can help restricting the candidate families of kernel functions. 
Table \ref{tab:kernel types} contains popular kernel families in literature and applications.
\begin{table}
    \caption{Kernel types}
    \resizebox{\textwidth}{!}{
    \begin{tabular}{lll}
        \toprule
       Kernel function & Equation & Hyperparameters \\
       \midrule
       Linear &  $k(x,y)=xy$ &   \\
       Polynomial &  $k(x,y)=(x^\intercal y+c)^d$ &   c, d\\
       Gaussian RBF &  $k(x,y)=\exp\left(-\frac{\|x-y\|_{2}^2}{2\l^2}\right)$ &   $l$\\
       Exponential RBF/Laplacian& $k(x,y)=\exp\left(-\frac{\|x-y\|_{2}}{l}\right)$ &   $l$\\
       Absolute Laplacian& $k(x,y)=\exp\left(-\frac{\|x-y\|_{1}}{l}\right)$ &   $l$\\
       Hyperbolic/Sigmoid Kernel &  $k(x,y2)=\tanh(\gamma x^\intercal y+r)$ &  $\gamma$, r \\
       Periodic &  $k(x,y)=\exp\left(\frac{-2 sin^2\left(\frac{\pi}{p}\|x-y\|_{2} \right)}{l^2}\right) $ &   $p, l$\\
       Chi-squared kernel &  $k(x,y)=\exp\left(-\gamma \sum\limits_i \frac{(x_i - y_i)^2}{x_i+y_i}\right) $ &   $\gamma$\\
        Cosine  &  $k(x,y)=\frac{xy^\intercal}{\|x\|_{2}\|y\|_{2}}$ &   \\
       Matérn  &  $k(x, y) =  \frac{1}{\Gamma(\nu)2^{\nu-1}}\Bigg(
        \frac{\sqrt{2\nu}}{l} \|x-y|\|_{2}
        \Bigg)^\nu Y_\nu\Bigg(
        \frac{\sqrt{2\nu}}{l} \|x-y|\|_{2}\Bigg)$ &  $l, \nu$ \\

@@

When $\nu=\frac{1}{2}$ the Matérn kernel corresponds to the Laplacian kernel. If the predictors are one dimensional, then the Absolute Laplacian and the Matern 0.5 are equivalent. %L2 norm=L1 norm
Furthermore, as $\nu \to \infty$ we have that the Matérn kernel converges to the Gaussian RBF kernel.
$k(x,y)=\exp\left(\frac{-\|x-y\|_{2}^2}{2l^2}\right)$

### 2024-05-31 thesis/input_files/probabilistic.tex
May 31st 2024

We conclude this chapter with reporting the same kind of Tables \ref{tab:kernel pinball comparison}, \ref{tab:kernel pinball comparison quantile-wise}, \ref{tab:kernel mae comparison} and Figure \ref{fig:kernel quantile regressors comparison} comparing various kernel functions.

% tables
\begin{table}[!h]
    \caption{Kernels comparison pinball loss for Melbourne data}
    \label{tab:kernel pinball comparison}
    \begin{center}
    \begin{tabular}{lll}
        \toprule
        & Kernel & Pinball loss
        \\
        \midrule
        & Absolute Laplacian &  10.136551 \\
        & Matern 0.5/Laplacian & 10.136551  \\
        & Matern 1.5 & 10.082668  \\
        & Matern 2.5 & 10.0494190    \\

>>>>

We conclude this chapter with reporting the same kind of Tables \ref{tab:kernel pinball comparison}, \ref{tab:kernel pinball comparison quantile-wise}, \ref{tab:kernel mae comparison} and Figure \ref{fig:kernel quantile regressors comparison} comparing various kernel functions, see Section \ref{kernel families}.

From these tables, we can conclude that the Matern and the Chi squared kernels are the top performers on the Melbourne weather data. Finally, the equaility in the scores of the Absolute Laplacian and the Matern 0.5 follows from the fact that regressors are one dimensional for the Melbourne weather data.
% tables
\begin{table}[!h]
    \caption{Kernels comparison pinball loss for Melbourne data}
    \label{tab:kernel pinball comparison}
    \begin{center}
    \begin{tabular}{lll}
        \toprule
        & Kernel & Pinball loss
        \\
        \midrule
        & Absolute Laplacian &  10.136551 \\
        & Matern 0.5/Laplacian & 10.136551  \\
        & Matern 1.5 & 10.082668  \\
        & Matern 2.5 & 10.0494190    \\
